---
title: "Assignment_week_5"
author: "Noam Ghenassia"
date: "2022-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

In this week's assignment, the objective is to observe how the optimal bandwidth $h_{opt}$ of a regression varies with the parameters of the distribution we are fitting. We will be computing the optimal $h_{opt}$ for a set of $n$ iid copies of a bivariate random vector $\{(X_i, Y_i)^T\}_{i \in 1,...,n}$, where $X \sim Beta(\beta_1, \beta_2)$ and $Y = sin(\frac{1}{X/3 + 0.1}) + \epsilon$, with $\epsilon \sim \mathcal{N}(0, \sigma^2)$. The following plot shows the function $m(x) = sin(\frac{1}{x/3 + 0.01})$ along with a sample drawn from our distribution (with $\beta_1 = \beta_2 = 1$ and $\sigma^2=0.1$ for the sake of clarity, even though for the rest of the assignment the value of $\sigma^2$ will be $1$) :

```{r m(x)}
library(ggplot2)
library(ggpubr)

b <- function(x){
  dbeta(x, shape1=1, shape2=1)
}

x_lin <- seq(0, 1, by=0.01)

m <- function(x){
  return(sin(1/(x/3 + 0.1)))
}

y_lin <- m(x_lin)
df1 <- data.frame(x=x_lin, y=y_lin)

n = 50
x <- rbeta(n, 1, 1)
e <- rnorm(n, 0, 0.1)
y <- m(x) + e
df2 <- data.frame(x=x, y=y)

ggplot(df1, aes(x, y)) + 
   geom_line() +
   geom_point(data = df2, color = "red")
```

The optimal values $h_{opt}$ are given by $h_{opt}(x) = n^{-1/5} \left[ \frac{\sigma^2 \int [K(z)]^2 dz}{[m''(x) \int z^2 K(z) dz]^2 f_X(x)} \right]^{1/5}$. We will now proceed to compute $h_{opt}(x)$, $x \in [0, 1]$ while varying the parameters $\beta_1$, $\beta_2$ of the distribution of $X$ and the number $n$ of samples drawn from it.

Here, we fix $\beta_1 = 1$ and plot $h_{opt}$ as a function of $x$ and $n$ for $\beta_2 = 1, 2, 3$:

```{r plots_1}

h_opt <- function(x, shape1, shape2, n) {
  ddxxm <- ((6*x + 1.8)*cos(3/(x+0.3)) - 9*sin(3/(x+0.3)))/((x+0.3)^4)
  denom <- (ddxxm * (1/12)) ^2 * dbeta(x, shape1, shape2)
  return(n^(-1/5) * (1/denom)^(1/5))
}


x <- seq(from=0, to=1, by=0.01)
df <- data.frame(x=x)

  
y100 <- h_opt(x, 1, 1, 100)
y500 <- h_opt(x, 1, 1, 500)
y1000 <- h_opt(x, 1, 1, 1000)
df_ <- data.frame(y_100_1_1=y100,
                 y_500_1_1=y500,
                 y_1000_1_1=y1000)
df <- cbind(df, df_)

y100 <- h_opt(x, 1, 2, 100)
y500 <- h_opt(x, 1, 2, 500)
y1000 <- h_opt(x, 1, 2, 1000)
df_ <- data.frame(y_100_1_2=y100,
                  y_500_1_2=y500,
                  y_1000_1_2=y1000)
df <- cbind(df, df_)

y100 <- h_opt(x, 1, 3, 100)
y500 <- h_opt(x, 1, 3, 500)
y1000 <- h_opt(x, 1, 3, 1000)
df_ <- data.frame(y_100_1_3=y100,
                  y_500_1_3=y500,
                  y_1000_1_3=y1000)
df <- cbind(df, df_)

plot1 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_1_1, colour="n=100")) + 
  geom_line(aes(y = y_500_1_1, colour="n=500")) +
  geom_line(aes(y = y_1000_1_1, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

plot2 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_1_2, colour="n=100")) + 
  geom_line(aes(y = y_500_1_2, colour="n=500")) +
  geom_line(aes(y = y_1000_1_2, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

plot3 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_1_3, colour="n=100")) + 
  geom_line(aes(y = y_500_1_3, colour="n=500")) +
  geom_line(aes(y = y_1000_1_3, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

ggarrange(plot1, plot2, plot3, ncol=1)
```

Here, we fix $\beta_1 = 2$ and plot $h_{opt}$ as a function of $x$ and $n$ for $\beta_2 = 1, 2, 3$:

```{r plots_2}
h_opt <- function(x, shape1, shape2, n) {
  ddxxm <- ((6*x + 1.8)*cos(3/(x+0.3)) - 9*sin(3/(x+0.3)))/((x+0.3)^4)
  denom <- (ddxxm * (1/12)) ^2 * dbeta(x, shape1, shape2)
  return(n^(-1/5) * (1/denom)^(1/5))
}


x <- seq(from=0, to=1, by=0.01)
df <- data.frame(x=x)

  
y100 <- h_opt(x, 2, 1, 100)
y500 <- h_opt(x, 2, 1, 500)
y1000 <- h_opt(x, 2, 1, 1000)
df_ <- data.frame(y_100_2_1=y100,
                 y_500_2_1=y500,
                 y_1000_2_1=y1000)
df <- cbind(df, df_)

y100 <- h_opt(x, 2, 2, 100)
y500 <- h_opt(x, 2, 2, 500)
y1000 <- h_opt(x, 2, 2, 1000)
df_ <- data.frame(y_100_2_2=y100,
                  y_500_2_2=y500,
                  y_1000_2_2=y1000)
df <- cbind(df, df_)

y100 <- h_opt(x, 2, 3, 100)
y500 <- h_opt(x, 2, 3, 500)
y1000 <- h_opt(x, 2, 3, 1000)
df_ <- data.frame(y_100_2_3=y100,
                  y_500_2_3=y500,
                  y_1000_2_3=y1000)
df <- cbind(df, df_)

plot1 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_2_1, colour="n=100")) + 
  geom_line(aes(y = y_500_2_1, colour="n=500")) +
  geom_line(aes(y = y_1000_2_1, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

plot2 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_2_2, colour="n=100")) + 
  geom_line(aes(y = y_500_2_2, colour="n=500")) +
  geom_line(aes(y = y_1000_2_2, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

plot3 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_2_3, colour="n=100")) + 
  geom_line(aes(y = y_500_2_3, colour="n=500")) +
  geom_line(aes(y = y_1000_2_3, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

ggarrange(plot1, plot2, plot3, ncol=1)
```

Here, we fix $\beta_1 = 3$ and plot $h_{opt}$ as a function of $x$ and $n$ for $\beta_2 = 1, 2, 3$:

```{r plots_3}
h_opt <- function(x, shape1, shape2, n) {
  ddxxm <- ((6*x + 1.8)*cos(3/(x+0.3)) - 9*sin(3/(x+0.3)))/((x+0.3)^4)
  denom <- (ddxxm * (1/12)) ^2 * dbeta(x, shape1, shape2)
  return(n^(-1/5) * (1/denom)^(1/5))
}


x <- seq(from=0, to=1, by=0.01)
df <- data.frame(x=x)

  
y100 <- h_opt(x, 3, 1, 100)
y500 <- h_opt(x, 3, 1, 500)
y1000 <- h_opt(x, 3, 1, 1000)
df_ <- data.frame(y_100_3_1=y100,
                 y_500_3_1=y500,
                 y_1000_3_1=y1000)
df <- cbind(df, df_)

y100 <- h_opt(x, 3, 2, 100)
y500 <- h_opt(x, 3, 2, 500)
y1000 <- h_opt(x, 3, 2, 1000)
df_ <- data.frame(y_100_3_2=y100,
                  y_500_3_2=y500,
                  y_1000_3_2=y1000)
df <- cbind(df, df_)

y100 <- h_opt(x, 3, 3, 100)
y500 <- h_opt(x, 3, 3, 500)
y1000 <- h_opt(x, 3, 3, 1000)
df_ <- data.frame(y_100_3_3=y100,
                  y_500_3_3=y500,
                  y_1000_3_3=y1000)
df <- cbind(df, df_)

plot1 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_3_1, colour="n=100")) + 
  geom_line(aes(y = y_500_3_1, colour="n=500")) +
  geom_line(aes(y = y_1000_3_1, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

plot2 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_3_2, colour="n=100")) + 
  geom_line(aes(y = y_500_3_2, colour="n=500")) +
  geom_line(aes(y = y_1000_3_2, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

plot3 <- ggplot(data=df, aes(x=x)) + 
  geom_line(aes(y = y_100_3_3, colour="n=100")) + 
  geom_line(aes(y = y_500_3_3, colour="n=500")) +
  geom_line(aes(y = y_1000_3_3, colour="n=1000")) +
  scale_colour_manual("", 
                      breaks = c("n=100", "n=500", "n=1000"),
                      values = c("red", "green", "blue"))

ggarrange(plot1, plot2, plot3, ncol=1)
```
