---
title: "Assignment_week_11"
author: "Noam Ghenassia"
date: "2022-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


## coverage of confidence intervals


```{r}
alpha <- 0.95
z_alpha <- qnorm(alpha)
n_obs <- 100
repeats <- 10^2
n_resamples <- 1000
```

In this assignment, we will compare the quality of asymptotic, studentized, non-studentized and sample-truth-scaled confidence intervals. To do that, we generate $10^3$ samples of 100 observations each, drawn from the exponential distribution with parameter $\lambda = 2$. For each sample, we then generate a confidence interval for the mean $\mathbb{E}[X] = 1/\lambda$ and verify whether it contains the true value $\mathbb{E}[X]=1/2$. the quality of the confidence interval can then be assessed from the coverage proportion, i.e., the fraction of CI's that conttain the true value. The results are as follows :

for the asymptotic distribution :

```{r assymptotic_coverage}
assymptotic_coverage_count <- 0
for (i in 1:repeats){
  X <- rexp(100, rate = 2)
  asymptotic <- mean(X) + z_alpha*sd(X)/sqrt(100)
  if (asymptotic >= 0.5){assymptotic_coverage_count <- assymptotic_coverage_count + 1}
}

print(assymptotic_coverage_count/repeats)
```

For the studentized CI :

```{r studentized_coverage}
studentized_coverage_count <- 0
for (i in 1:repeats){
  X <- rexp(100, rate = 2)
  boot <- sample(1:n_obs, size = 100*n_resamples, replace = T)
  boot_data <- matrix(X[boot], ncol = n_resamples)
  T_star <- c()
  for (b in 1:n_resamples){
    tstar <- sqrt(n_obs)*(mean(boot_data[,b])-mean(X))/sd(boot_data[,b])
    T_star[b] <- append(T_star, tstar)
  }
  studentized <- mean(X) + quantile(T_star, 0.95)*sd(X)/sqrt(100)
  if (studentized >= 0.5){studentized_coverage_count <- studentized_coverage_count + 1}
}

print(studentized_coverage_count/repeats)
```

For the non-studentized CI :

```{r non_studentized_coverage}
non_studentized_coverage_count <- 0
repeat_count <- 0
for (i in 1:repeats){
  repeat_count <- repeat_count +1
  X <- rexp(100, rate = 2)
  boot <- sample(1:n_obs, size = 100*n_resamples, replace = T)
  boot_data <- matrix(X[boot], ncol = n_resamples)
  T_star <- c()
  for (b in 1:n_resamples){
    tstar <- sqrt(n_obs)*(mean(boot_data[,b])-mean(X))
    T_star[b] <- append(T_star, tstar)
  }
  non_studentized <- mean(X) + quantile(T_star, 0.95)*sd(X)/sqrt(100)
  if (non_studentized >= 0.5){non_studentized_coverage_count <- non_studentized_coverage_count + 1}
}

print(non_studentized_coverage_count/repeats)
```

for the sample-truth-scaled :

```{r sample_truth_scaled_coverage}
sample_truth_scaled_coverage_count <- 0
for (i in 1:repeats){
  X <- rexp(100, rate = 2)
  boot <- sample(1:100, size = 100*n_resamples, replace = T)
  boot_data <- matrix(X[boot], ncol = n_resamples)
  T_star <- c()
  for (b in 1:n_resamples){
    tstar <- sqrt(n_obs)*(mean(boot_data[,b])-mean(X))/sd(X)
    T_star[b] <- append(T_star, tstar)
  }
  sample_truth_scaled <- mean(X) + quantile(T_star, 0.95)*sd(X)/sqrt(100)
  if (sample_truth_scaled >= 0.5){sample_truth_scaled_coverage_count <- sample_truth_scaled_coverage_count + 1}
}

print(sample_truth_scaled_coverage_count/repeats)
```

